---
title: "GPS Footprinting"
output:
  pdf_document: default
  rmarkdown::github_document
---

This is the *GPS Footprinting project*. It is inspired by [Finn et al. Nat Neuro  (2015)](https://www.nature.com/articles/nn.4135) and [Kaufmann et al. Nat Neuro (2015)](https://www.nature.com/articles/s41593-019-0471-7).

GPS data preprocessing performed according to [Ian Barnett's imputation algorithm](https://github.com/ianjamesbarnett/SmartphoneSensorPipeline), originally published in [Biostatistics (2020)](https://academic.oup.com/biostatistics/article-abstract/21/2/e98/5145908).

### Load Required Packages 
```{r load_lib, message=FALSE}
require(ggplot2)
require(summarytools)
require(cowplot)
source('~/Documents/GitHub/SmartphoneSensorPipeline/Extra/plotting_functions.R')
```

### Setup the paths 
```{r def_paths}
project_path = "~/Documents/xia_gps/"
data_path = file.path(project_path,"beiwe_output_043020")

gps_df_path = file.path(data_path,"Processed_Data/Group/feature_matrix.txt")
```

### Read in data
```{r read_gps}
#colClasses=c('character','character',rep("NULL",94),rep("numeric",15),rep("NULL",16))
gps_df = read.table(gps_df_path,header = T, dec = ",", )[,c(1,2,97:111)]
#gps_df = read.csv2(gps_df_path,header = T, sep= "", na.strings ="",colClasses= colClasses)
head(gps_df)
```

The current data has `r length(unique(gps_df$IID))` subjects, consisting of `r dim(gps_df)[1]` total days, and `r length(colnames(gps_df))-2` GPS features.

```{r fig.width=10, fig.height=4, echo=FALSE}
minMiss_histplot(gps_df,200, "All Data")
```


```{r define_col_types}
#Define the column types
gps_df$Date = as.Date(gps_df$Date)
gps_df[,3:dim(gps_df)[2]] = apply(gps_df[,3:dim(gps_df)[2]], 2, function(x) as.numeric(x))
```


### Exclude Data
#### 1.  remove 1st and last days
```{r 1st_last_days}
# loop through each subj to remove 1st and last days of gps data
gps_df_clean = data.frame() #initiate a df
for (subj in unique(gps_df$IID)){ #loop through each subj
  gps_df_subj <- subset(gps_df, IID == subj) #get gps_df per subject
  gps_df_subj <- gps_df_subj[2:(dim(gps_df_subj)[1]-1),] #remove the 1st and last days
  gps_df_clean <- rbind(gps_df_clean,gps_df_subj) #combine all subjs
}
```

This step removed `r dim(gps_df)[1] - dim(gps_df_clean)[1]` days. Now, dataset has `r length(unique(gps_df_clean$IID))` subjects, consisting of `r dim(gps_df_clean)[1]` total days, and `r length(colnames(gps_df_clean))-2` GPS features.

```{r fig.width=10, fig.height=4, echo=FALSE}
minMiss_histplot(gps_df_clean,200, "After Removing 1st and Last Days")
```

#### 2.  remove days with all mins missing
```{r results='asis'}
gps_df_clean2 = subset(gps_df_clean, MinsMissing < 1440)
```

This step removed `r dim(gps_df_clean)[1] - dim(gps_df_clean2)[1]` days. Now, dataset has `r length(unique(gps_df_clean2$IID))` subjects, consisting of `r dim(gps_df_clean2)[1]` total days, and `r length(colnames(gps_df_clean2))-2` GPS features.

```{r fig.width=10, fig.height=4, echo=FALSE}
minMiss_histplot(gps_df_clean2,200, "After Removing All-Day-Missing ")
```

```{r fig.width=10, fig.height=4, echo=FALSE}
minMiss_histplot(subset(gps_df_clean2,MinsMissing>=1296),200, "Zoom In 1296-1440")
```
